---
title: "mushrooms_assignment3"
author: "Ravi Kothari"
date: "July 31, 2016"
output: html_document
---
## Task
Your task is to study the dataset and the associated description of the data (i.e. "data dictionary"). You may need
to look around a bit, but it's there! You should take the data, and create a data frame with a subset of the columns
(and if you like rows) in the dataset. You should include the column that indicates edible or poisonous and three or
four other columns. You should also add meaningful column names and replace the abbreviations used in the
data-for example, in the appropriate column, "e" might become "edible." Your deliverable is the R code to
perform these transformation tasks.

## Solution
These are the steps we performed to get the resulted dataset:

1. loaded the dataset from the URL https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data

```{r load_dataset}
dataURL <- "https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.data"
mushrooms <- read.table(file = dataURL, header = FALSE, sep = ",")
```

2. Subsetting the data to include first 50 rows and change the column names to meaningful names.

```{r rename_column_name, results="hide"}
str(mushrooms)
sub.mushrooms <- mushrooms[1:50, ]
names(sub.mushrooms) <- c("Classes(edible/poisnous)", "cap-shape", "cap-surface", "cap-color", "bruises?", "odor", "gill-attachment", "gill-spacing", "gill-size", "gill-color","stalk-shape", "stalk-root", "stalk-surface-above-ring", "stalk-surface-below-ring", "stalk-color-above-ring", "stalk-color-below-ring", "veil-type", "veil-color", "ring-number", "ring-type", "spore-print-color", "population", "habitat")
sub.mushrooms
```

3. Filtering data to only include 5 columns and then renaming factors levels of these column to more meaningful name based on the data description in the dataset on URL https://archive.ics.uci.edu/ml/machine-learning-databases/mushroom/agaricus-lepiota.names

```{r renaming factors level}
sample.Mushroom <- sub.mushrooms[ , c("Classes(edible/poisnous)", "cap-shape", "cap-surface", "cap-color", "population")]
summary(sample.Mushroom)
levels(sample.Mushroom$`Classes(edible/poisnous)`) <- c("edible", "poisnous")
levels(sample.Mushroom$`cap-shape`) <- c("Bell", "Conical", "Flat", "Knobbed", "Sunken", "Convex")
levels(sample.Mushroom$`cap-surface`) <- c("Fibrous", "Grooves", "Smooth", "Scaly")
#cap-color:brown=n,buff=b,cinnamon=c,gray=g,green=r,pink=p,purple=u,red=e,white=w,yellow=y
levels(sample.Mushroom$`cap-color`) <- c("buff","cinnamon","red","gray","brown","pink","green","purple","white","yellow")
levels(sample.Mushroom$`population`) <- c("abundant","clustered","numerous","scattered","several","solitary")
#population:abundant=a,clustered=c,numerous=n,scattered=s,several=v,solitary=y
sample.Mushroom
```


