---
title: "Assignment 2"
author: "Ravi Kothari"
date: "July 24, 2016"
output: html_document
---
## Task

R has built in functions for choose() and factorial(). Your task is to write your own variation of the choose()
function, that does not use either of these functions. 


## Solution
The two functions below calculate factorial and combination for any two numbers n and k, with n being the number of items available to you and k the number of items you can choose at one time (where order does not matter).

R factorial function:

```{r factorial}
factorial <- function(x){
  y <- 1
  for(i in 1:x){
    y <-y*((1:x)[i])
    }
  return (y)
  }
```

R combination function:
```{r combination}
combination <- function(n,k){
  if(n < 0 || k > n){
    "Please enter a positive integer where n is greater than k"
  } else {
    numerator <- factorial(n)
    denominator <- (factorial(k)*factorial(n-k))
    answer <- numerator/denominator
    return(answer)
  }
  
}
```

## Test Cases

```{r test_cases}
combination(5,3)
combination(10,3)
combination(3,10)
combination(-3,-10)
```

